---
title: "Our Lego Sets"
author: "Stephen Kaluzny"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

```{r load_packages, echo=FALSE}
suppressPackageStartupMessages(library("here"))
suppressPackageStartupMessages(library("dplyr"))
suppressPackageStartupMessages(library("ggplot2"))
```

## The Data

Read the data from the LibreOffice spreadsheet.

```{r read_data}
cache_data <- here::here(file.path("data", "our_lego.rds"))
if(file.exists(cache_data)) {
  d <- readRDS(cache_data)
} else {
  d <- readODS::read_ods(here::here(file.path("data", "our_lego.ods")),
    sheet="LEGO")
  saveRDS(d, cache_data)
}
```

Select only our Lego, omitting the Lego we have given away or are
planning to give away.

```{r only_our_lego}
indxGive_Away <- grep("Give Away", d[[1]])
stopifnot(length(indxGive_Away) == 1)
d <- d[seq_len(indxGive_Away - 1), ]
```

Clean up the data.

```{r clean}
# Drop all blank rows
indxBlank <- which(apply(d, 1, function(x) sum(is.na(x))) == ncol(d))
d <- d[-indxBlank, ]
```

```{r convert_date}
d[["Date"]] <- as.Date(d[["Date"]], format="%m/%d/%y")
d[["PurchaseYear"]] <- lubridate::year(d[["Date"]])
```

```{r convert_cost}
d[["Cost"]] <- as.numeric(d[["Cost"]])
stopifnot(all(d[["Cost"]] >=0, na.rm=TRUE))
```

```{r names}
names(d)
```

```{r class}
sapply(d, class)
```

## Summary

The resulting dataset has `r nrow(d)` observations.

Total number of pieces is `r format(sum(d[["Pieces"]], na.rm=TRUE), digits=6)`.

The first Lego was acquired `r min(d[["Date"]], na.rm=TRUE)`.

## Plots

```{r sets_by_year}
d %>% group_by(PurchaseYear) %>%
  summarize(N = n()) %>%
  ggplot(aes(x=PurchaseYear, y=N)) +
    geom_line()
```

```{r pieces_by_year}
d %>% group_by(PurchaseYear) %>%
  summarize(PiecesYear = sum(Pieces, na.rm=TRUE)) %>%
  ggplot(aes(x=PurchaseYear, y=PiecesYear)) +
    geom_line()
```

```{r cost_by_year}
d %>% group_by(PurchaseYear) %>%
  summarise(CostYear = sum(Cost, na.rm=TRUE)) %>%
  filter(PurchaseYear > 2015) %>%
  # knitr::kable() %>%
  ggplot(aes(x=PurchaseYear, y=CostYear)) +
    geom_line()
```

## Appendix
This is a reproducible document.
All code are available at https://github.com/spkaluzny/legolog.
The data is not publicly available.

```{r sessionInfo}
sessionInfo()
```
