---
title: "Our Lego Sets"
author: "Stephen Kaluzny"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

```{r load_packages, echo=FALSE}
# suppressPackageStartupMessages(library("here"))
suppressPackageStartupMessages(library("dplyr"))
suppressPackageStartupMessages(library("ggplot2"))
```

## The Data

Read the data from the LibreOffice spreadsheet.

```{r read_data}
cache_data <- here::here(file.path("data", "our_lego.rds"))
if(file.exists(cache_data)) {
  d <- readRDS(cache_data)
} else {
  d <- readODS::read_ods(here::here(file.path("data", "our_lego.ods")),
    sheet="LEGO")
  saveRDS(d, cache_data)
}
```

Select only our Lego, omitting the Lego we have given away or are
planning to give away.

```{r only_our_lego}
indxGive_Away <- grep("Give Away", d[[1]])
stopifnot(length(indxGive_Away) == 1)
d <- d[seq_len(indxGive_Away - 1), ]
```

Clean up the data.

Remove all blank lines:
```{r clean}
indxBlank <- which(apply(d, 1, function(x) sum(is.na(x))) == ncol(d))
d <- d[-indxBlank, ]
```

Convert date, compute `PurchaseYear` and sort by date:
```{r convert_date}
d[["Date"]] <- as.Date(d[["Date"]], format="%m/%d/%y")
d[["PurchaseYear"]] <- lubridate::year(d[["Date"]])
o <- order(d[["Date"]])
d <- d[o, ]
```

Clean up `Cost`:
```{r convert_cost}
d[["Cost"]] <- as.numeric(d[["Cost"]])
stopifnot(all(d[["Cost"]] >=0, na.rm=TRUE))
```

```{r names}
names(d)
```

```{r class}
sapply(d, class)
```

## Summary

The resulting dataset has `r nrow(d)` observations.

Total number of pieces is `r format(sum(d[["Pieces"]], na.rm=TRUE), digits=6)`.

The first Lego was acquired `r min(d[["Date"]], na.rm=TRUE)`.

## Plots

### Sets by Year

```{r sets_by_year}
d %>% filter(!is.na(PurchaseYear)) %>%
  group_by(PurchaseYear) %>%
  summarize(N = n()) %>%
  ggplot(aes(x=PurchaseYear, y=N)) +
    geom_line() +
    geom_point() +
    ggtitle("Number of Sets Purchased by Year")
```

```{r cumsum_sets}
d %>%
  filter(!is.na(Date)) %>%
  mutate(TotalSets = seq.int(length(Date))) %>%
  ggplot(aes(x=Date, y=TotalSets)) +
    geom_point(cex=0.5) +
    ggtitle("Cumulative Number of Sets by Date")
```

### Pieces by Year

```{r pieces_by_year}
d %>% filter(!is.na(PurchaseYear)) %>%
  group_by(PurchaseYear) %>%
  summarize(PiecesYear = sum(Pieces, na.rm=TRUE)) %>%
  ggplot(aes(x=PurchaseYear, y=PiecesYear)) +
    geom_line() +
    geom_point() +
    ggtitle("Number of Pieces Purchased by Year")
```

```{r cumsum_pieces}
d %>% 
  filter(!is.na(Pieces), !is.na(Date)) %>%
  mutate(CumPieces = cumsum(Pieces)) %>%
  ggplot(aes(x=Date, y=CumPieces)) +
    geom_point(cex=0.5) +
    ggtitle("Cumulative Pieces Over Time")
```

### Spending by Year

```{r spending_by_year}
d %>% filter(!is.na(PurchaseYear)) %>%
  group_by(PurchaseYear) %>%
  summarise(CostYear = sum(Cost, na.rm=TRUE)) %>%
  # filter(PurchaseYear > 2015) %>%
  # knitr::kable() %>%
  ggplot(aes(x=PurchaseYear, y=CostYear)) +
    geom_line() +
    geom_point() +
    ggtitle("Lego Spending by Year")
```

```{r cumsum_spending}
d %>%
  filter(!is.na(Cost), !is.na(Date)) %>%
  mutate(CumCost = cumsum(Cost)) %>%
  ggplot(aes(x=Date, y=CumCost)) +
    geom_point(cex=0.5) +
    ggtitle("Cumulative Lego Spending Over Time")
```

## Appendix
This is a reproducible document.
All code are available at https://github.com/spkaluzny/legolog.
The data is not publicly available.

```{r sessionInfo}
sessionInfo()
```
